---
title: "In-class_Ex07: Emerging Hot Spot Analysis: sfdep methods"
date: "20 February 2023"
date-modified: "r Sys.Date()"
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

## Overview

Emerging Hot Spot Analysis (EHSA) is a spatio-temporal analysis method for revealing and describing how hot spot and cold spot areas evolve over time. The analysis consist of four main steps:

-   Building a space-time cube,
-   Calculating Getis-Ord local Gi\* statistic for each bin by using an FDR correction,
-   Evaluating these hot and cold spot trends by using Mann-Kendall trend test,
-   Categorising each study area location by referring to the resultant trend z-score and p-value for each location with data, and with the hot spot z-score and p-value for each bin.

## Getting Started

### Installing and Loading the R Packages

As usual, `p_load()` of pacman package will be used to check if the necessary packages have been installed in R, if yes, load the packages on R environment.

Five R packages are need for this in-class exercise, they are: sf, sfdep, tmap, and tidyverse.

```{r}
pacman::p_load(sf, sfdep, tmap, plotly, tidyverse, zoo)
```

### The Data

For the purpose of this in-class exercise, the Hunan data sets will be used. There are two data sets in this use case, they are:

-   Hunan, a geospatial data set in ESRI shapefile format, and
-   Hunan_GDPPC, an attribute data set in csv format. Before getting started, reveal the content of *Hunan_GDPPC.csv* by using Notepad and MS Excel.

### Importing geospatial data

In the code chunk below, st_read() of sf package is used to import *Hunan* shapefile into R.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Importing attribute table

In the code chunk below, `read_csv()` of readr is used to import *Hunan_GDPPC.csv* into R.

```{r}
GDPPC <- read_csv("data/aspatial/Hunan_GDPPC.csv")
```

## Creating a Time Series Cube

In the code chunk below, `spacetime()` of sfdep ised used to create an spatio-temporal cube.

```{r}
GDPPC_st <- spacetime(GDPPC, hunan,
                      .loc_col = "County",
                      .time_col = "Year")
```

Next, `is_spacetime_cube()` of sfdep package will be used to verify if GDPPC_st is indeed an space-time cube object.

```{r}
is_spacetime_cube(GDPPC_st)
```
